### 2020-04-05 thesis/content/chapter-evaluation.tex
added thesis structure
started with ltag definition + various minor fixes
continued with spatial gcnn lta relation
continued with k-gnn limits section
various fixes + started with evaluation chapter

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Introduction}%
\label{sec:intro}

\pagenumbering{arabic}			% arabic page numbering
\setcounter{page}{1}			% set page counter

\section{Motivation}%
\label{sec:intro:motivation}

\section{Goals}%
\label{sec:intro:goals}

\section{Structure}%
\label{sec:intro:structure}

\paragraph{\Cref{sec:conclusion}: \nameref{sec:conclusion}}

>>>>

%!TEX root = ../main.tex
% chktex-file 46
\chapter{Evaluation}%
\label{sec:eval}

We will now empirically evaluate the ideas presented in the previous chapters.
In \cref{sec:ltag} the relation between \ac{lta} and existing \ac{gcr} approaches was formally analyzed.
There we saw that the \acs{lta}-formulations of existing approaches mostly use static decomposition functions, e.g.\ \acs{bfs}-subtree decompositions.
Motivated by the idea of dynamically learning decompositions via edge filters, we proposed the novel 2-\acs{wl}-\acs{gnn} in the last chapter.
Note however that a 2-\acs{wl}-\acs{gnn} by itself is not more ``\acs{lta}-like'' than other \acp{gnn}.

In order to empirically evaluate the theoretical results of this thesis, we therefore differentiate between two independent evaluation goals:
\begin{enumerate}[label={\textbf{\arabic*.}}]
	\item \textbf{Comparison of 2-\acs{wl}-\acsp{gnn} with other \ac{gnn} methods:}
		Even though it was motivated by \ac{lta}, the 2-\acs{wl} convolution layer should be compared with other convolution approaches in a non-\acs{lta} configuration.
	\item \textbf{Comparison of \acs{lta}-like \ac{gcr} methods with non-\acs{lta} methods:}
\end{enumerate}

\section{Experimental Setup}%
\label{sec:eval:setup}

\section{Synthetic Datasets}%
\label{sec:eval:synthetic}

\section{Real-World Datasets}%
\label{sec:eval:real}

### 2020-04-05 thesis/content/chapter-ltd.tex
various fixes + started with evaluation chapter

		\item\label{inv:ltd:wl2-simulation:indicator}

@@

		\item\label{inv:ltd:wl2-simulation:feature}

@@

		\item\label{inv:ltd:wl2-simulation:weight}

@@

	This means that the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.

@@

	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve the invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:weight}.

@@

	Using invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
	\begin{align*}
		\varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \coloneqq \begin{cases}
			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
			\texttt{nil} & \text{else}
		\end{cases}
		\text{.}
	\end{align*}
	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
\end{proof}
\begin{cor}\label{cor:ltd:wl2-gnn-wl1-power}
	2-\acs{wl}-\acsp{gnn} have at least the same discriminative power as 1-\acs{wl}.
\end{cor}
\begin{proof}
	The corollary directly follows from the fact that 2-\acs{wl}-\acsp{gnn} can simulate \acp{gin} by \cref{thm:ltd:wl2-simulation} and the fact that \acp{gin} have the same discriminative power as 1-\acs{wl} by \cref{prop:related:gcnn-wl1-limit} (see \cpageref{prop:related:gcnn-wl1-limit}) because they use injective vertex neighborhood hashing functions~\cite{Xu2018}.
\end{proof}

@@

		\rowfont{\color{t_blue}}$1$ & $1$ & $1$ \\
		\rowfont{\color{t_blue}}$1$ & $4$ & $4$ \\
		\rowfont{\color{t_blue}}$1$ & $5$ & $5$ \\
		\rowfont{\color{t_blue}}$2$ & $2$ & $2$ \\
		\rowfont{\color{t_blue}}$2$ & $4$ & $4$ \\
		\rowfont{\color{t_blue}}$2$ & $6$ & $6$ \\
		\rowfont{\color{t_blue}}$3$ & $3$ & $3$ \\
		\rowfont{\color{t_blue}}$3$ & $5$ & $5$ \\
		\rowfont{\color{t_blue}}$3$ & $6$ & $6$ \\
		\rowfont{\color{t_red}}$4$ & $1$ & $4$ \\
		\rowfont{\color{t_red}}$4$ & $4$ & $2$ \\
		\rowfont{\color{t_red}}$4$ & $5$ & $6$ \\
		\rowfont{\color{t_red}}$5$ & $1$ & $5$ \\
		\rowfont{\color{t_red}}$5$ & $5$ & $3$ \\
		\rowfont{\color{t_red}}$5$ & $4$ & $6$ \\
		\rowfont{\color{t_red}}$6$ & $2$ & $6$ \\
		\rowfont{\color{t_red}}$6$ & $6$ & $3$ \\
		\rowfont{\color{t_red}}$6$ & $4$ & $5$ \\

@@

		\rowfont{\color{t_blue}}$7$ & $7$ & $7$ \\
		\rowfont{\color{t_blue}}$7$ & $9$ & $9$ \\
		\rowfont{\color{t_blue}}$8$ & $8$ & $8$ \\
		\rowfont{\color{t_blue}}$8$ & $9$ & $9$ \\
		\rowfont{\color{t_red}}$9$ & $7$ & $9$ \\
		\rowfont{\color{t_red}}$9$ & $9$ & $8$

@@

Using the $\mathit{gather}$ and $\mathit{scatter}_{\Sigma}$ operator, the 2-\acs{wl} convolution operator from \cref{defn:ltd:wl2-conv-step} can be computed via the following algorithm:
\begin{algorithm}[H]
	\caption{Parallel Implementation of a 2-\acs{wl} Convolution Layer $S^{(t)}$}\label{algo:ltd:wl2-conv}
	\begin{algorithmic}[1]
		\Function{$S^{(t)}$}{$Z^{(t-1)} \in \mathbb{R}^{m \times d^{(t-1)}}, R \in {[m]}^{\gamma \times 3}$}
			\State{$Z_{\mathrm{L}} \coloneqq Z^{(t-1)} W_{\mathrm{L}}^{(t)}$}
			\Comment{Matrix multiply: $\mathbb{R}^{m \times d^{(t-1)}} \to \mathbb{R}^{m \times d^{(t)}}$}
			\State{$Z_{\mathrm{F}} \coloneqq Z^{(t-1)} W_{\mathrm{F}}^{(t)}$}
			\State{$Z_{\Gamma} \coloneqq Z^{(t-1)} W_{\Gamma}^{(t)}$}
			\State{$X_{\Gamma, 1} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 1})$}
			\Comment{Gather: $\mathbb{R}^{m \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{\gamma \times d^{(t)}}$}
			\State{$X_{\Gamma, 2} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 2})$}
			\State{$X_{\Gamma} \coloneqq \sigma_{\Gamma}\left(X_{\Gamma, 1} + X_{\Gamma, 2}\right)$}
			\Comment{Element-wise operations}
			\State{$Z_{\Sigma\Gamma} \coloneqq \mathit{scatter}_{\Sigma}(X_{\Gamma}, R_{\mathrm{L}})$}
			\Comment{Scatter: $\mathbb{R}^{\gamma \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{m \times d^{(t)}}$}
			\State{$Z^{(t)} \coloneqq \sigma\left(Z_{\mathrm{L}} + Z_{\mathrm{F}} \odot Z_{\Sigma\Gamma} \right)$}
			\Comment{Element-wise operations}
			\State{\Return{$Z^{(t)}$}}
		\EndFunction{}
	\end{algorithmic}
\end{algorithm}
All operations in this implementation scale well on parallel computing hardware and are supported by common \ac{ml} libraries like \citetitle{TF}~\cite{Abadi2015}\cite{TF} or \citetitle{PyT}~\cite{Paszke2019}\cite{PyT}.
2-\acs{wl} convolutions can therefore be easily integrated into existing \ac{ml} pipelines.
This concludes our description of 2-\acs{wl}-\acsp{gnn}.

>>>>

		\item\label[inv]{inv:ltd:wl2-simulation:indicator}

@@

		\item\label[inv]{inv:ltd:wl2-simulation:feature}

@@

		\item\label[inv]{inv:ltd:wl2-simulation:weight}

@@

	This means that \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight} are satisfied after $S^{(t, 2)}$.

@@

	Via the universal approximation theorem~\cite{Hornik1991} we can therefore use $S^{(t, 2+1)}, \dots, S^{(t, 2+K)}$ to simulate the $K$ layers of $\mathrm{\acs*{mlp}}^{(t)}$ without changing the first and last dimension of each feature vector to preserve \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:weight}.

@@

	Using \cref{inv:ltd:wl2-simulation:indicator,inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
	\begin{align*}
		\varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \coloneqq \begin{cases}
			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
			\texttt{nil} & \text{else}
		\end{cases}
		\text{.}
	\end{align*}
	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
\end{proof}
\begin{cor}\label{cor:ltd:wl2-gnn-wl1-power}
	2-\acs{wl}-\acsp{gnn} have at least the same discriminative power as 1-\acs{wl}.
\end{cor}
\begin{proof}
	The corollary directly follows from the fact that 2-\acs{wl}-\acsp{gnn} can simulate \acp{gin} by \cref{thm:ltd:wl2-simulation} and the fact that \acp{gin} have the same discriminative power as 1-\acs{wl} by \cref{prop:related:gcnn-wl1-limit} (see \cpageref{prop:related:gcnn-wl1-limit}) because they use injective vertex neighborhood hashing functions~\cite{Xu2018}.
\end{proof}

@@

		\rowfont{\color{t_blue}}$(1,$ & $1,$ & $1)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(1,$ & $4,$ & $4)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(1,$ & $5,$ & $5)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(2,$ & $2,$ & $2)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(2,$ & $4,$ & $4)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(2,$ & $6,$ & $6)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(3,$ & $3,$ & $3)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(3,$ & $5,$ & $5)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(3,$ & $6,$ & $6)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(4,$ & $1,$ & $4)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(4,$ & $4,$ & $2)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(4,$ & $5,$ & $6)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(5,$ & $1,$ & $5)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(5,$ & $5,$ & $3)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(5,$ & $4,$ & $6)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(6,$ & $2,$ & $6)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(6,$ & $6,$ & $3)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(6,$ & $4,$ & $5)$ \\ % chktex 40

@@

		\rowfont{\color{t_blue}}$(7,$ & $7,$ & $7)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(7,$ & $9,$ & $9)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(8,$ & $8,$ & $8)$ \\ % chktex 40
		\rowfont{\color{t_blue}}$(8,$ & $9,$ & $9)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(9,$ & $7,$ & $9)$ \\ % chktex 40
		\rowfont{\color{t_red}}$(9,$ & $9,$ & $8)$ % chktex 40

@@

Using the $\mathit{gather}$ and $\mathit{scatter}_{\Sigma}$ operator, the 2-\acs{wl} convolution operator from \cref{defn:ltd:wl2-conv-step} can be computed via the following algorithm:
\begin{algorithm}[H]
	\caption{Parallel Implementation of a 2-\acs{wl} Convolution Layer $S^{(t)}$}\label{algo:ltd:wl2-conv}
	\begin{algorithmic}[1]
		\Function{$S^{(t)}$}{$Z^{(t-1)} \in \mathbb{R}^{m \times d^{(t-1)}}, R \in {[m]}^{\gamma \times 3}$}
			\State{$Z_{\mathrm{L}} \coloneqq Z^{(t-1)} W_{\mathrm{L}}^{(t)}$}
			\Comment{Matrix multiply: $\mathbb{R}^{m \times d^{(t-1)}} \to \mathbb{R}^{m \times d^{(t)}}$}
			\State{$Z_{\mathrm{F}} \coloneqq Z^{(t-1)} W_{\mathrm{F}}^{(t)}$}
			\State{$Z_{\Gamma} \coloneqq Z^{(t-1)} W_{\Gamma}^{(t)}$}
			\State{$X_{\Gamma, 1} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 1})$}
			\Comment{Gather: $\mathbb{R}^{m \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{\gamma \times d^{(t)}}$}
			\State{$X_{\Gamma, 2} \coloneqq \mathit{gather}(Z_{\Gamma}, R_{\Gamma, 2})$}
			\State{$X_{\Gamma} \coloneqq \sigma_{\Gamma}\left(X_{\Gamma, 1} + X_{\Gamma, 2}\right)$}
			\Comment{Element-wise operations}
			\State{$Z_{\Sigma\Gamma} \coloneqq \mathit{scatter}_{\Sigma}(X_{\Gamma}, R_{\mathrm{L}})$}
			\Comment{Scatter: $\mathbb{R}^{\gamma \times d^{(t)}} \times {[m]}^{\gamma} \to \mathbb{R}^{m \times d^{(t)}}$}
			\State{$Z^{(t)} \coloneqq \sigma\left(Z_{\mathrm{L}} + Z_{\mathrm{F}} \odot Z_{\Sigma\Gamma} \right)$}
			\Comment{Element-wise operations}
			\State{\Return{$Z^{(t)}$}}
		\EndFunction{}
	\end{algorithmic}
\end{algorithm}
All operations in this implementation scale well on parallel computing hardware, they are differentiable and are supported by common \ac{ml} libraries like \citetitle{TF}~\cite{Abadi2015}\cite{TF} and \citetitle{PyT}~\cite{Paszke2019}\cite{PyT}.
2-\acs{wl} convolutions can therefore be easily integrated into existing \ac{ml} tooling and be optimized via well-known gradient-based methods like \textit{Adam}~\cite{Kingma2015}.
This concludes our description of 2-\acs{wl}-\acsp{gnn}.
